\chapter{Third-Party Code and Libraries}

\section{Plotly.js}
Plotly.js is an open source charting and data visualisation library written in Javascript. Plotly.js is released under the MIT license. It was used in this project to provide the means of visualising data in graphical format.

\section{jQuery}
jQuery is a convenience library built to simplify development in Javascript, providing shorthand approaches to common tasks. jQuery is released under the MIT license. jQuery was used in this project to streamline the development of Javascript components and provide a more readable and maintainable solution to complex functionality such as asynchronous updates between the front end and the server.
\section{Bootstrap}
Bootstrap is a Javascript framework that provides simple implementations of common front end design and styling using combinations of HTML, Javascript and CSS. Bootstrap provides many convenience functions and plugins to support quick development of front end web pages and interactions within them. Bootstrap is released under the MIT license. Bootstrap was used in this project to provide a grid system layout for the pages in the system and generic CSS to provide a simple yet elegant look and feel. 
\section{OpenCSV}
OpenCSV is an open source CSV file parser library written for Java. OpenCSV is released under the Apache Software License, Version 2.0. OpenCSV was used in this project to provide a lightweight and simple method of parsing CSV files into a specific representation within the system.
\section{Spring}
The Spring framework is a large scale Java platform providing architectural and infrastructure handling with the aim of allowing developers to focus on application development. Spring provides dependency injection and inversion of control based environments. The Spring framework is released under the Apache Software License, Version 2.0. The Spring framework provides the underlying structure and architecture on which the delivered system is built.
\section{Maven}
Maven is a build tool providing the facility to automate the building of Java projects. Maven also provides a simple way to define and inject library dependencies into a project. Maven is released under the Apache Software License, Version 2.0. Maven was used as the build automation tool for this project.
\section{Tomcat}
Apache Tomcat is an open source web server implementation for Java. Tomcat is released under the Apache Software License, Version 2.0. Tomcat is the web server technology that run the delivered system.